---
title: "Ordination Analysis"
---
# Setup
```{r}
library(phyloseq)
library(viridis)
library(tidyverse)
library(ggpubr)
```
# Import filtered data
```{r}
filtPhyseq <- readRDS("../data/PHYLOSEQ_DATA/phyloseq_filtered_rare.rds")
filtPhyseq
summary(sample_sums(filtPhyseq))
```
# Import filtered data
```{r}
filtPhyseqRare <- readRDS("../data/PHYLOSEQ_DATA/phyloseq_filtered_rare.rds")
filtPhyseqRare
summary(sample_sums(filtPhyseqRare))
```

# Transform dataset
```{r}
filtPhyseqProp <- transform_sample_counts(filtPhyseq, function(x){100 * x / sum(x)})
filtPhyseqLog <- transform_sample_counts(filtPhyseq, function(x) log(1 + x))
```

# Transform rarified dataset
```{r}
filtPhyseqRareProp <- transform_sample_counts(filtPhyseqRare, function(x){100 * x / sum(x)})
filtPhyseqRareLog <- transform_sample_counts(filtPhyseqRare, function(x) log(1 + x))
```

# Ordination - all samples

## Method Exploration
Run through some quick ordination plots to look at different ordinations methods
```{r}
dist = "wunifrac"
ord_meths = c("DCA", "CCA", "RDA", "DPCoA", "NMDS", "MDS")
plist = plyr::llply(as.list(ord_meths), function(i, physeq, dist){
        ordi = ordinate(physeq, method=i, distance=dist)
        plot_ordination(physeq, ordi, "samples", color="DIAG_CRS")
}, filtPhyseqRareProp, dist)
```
```{r}
names(plist) <- ord_meths
```
```{r}
pdataframe = plyr::ldply(plist, function(x){
    df = x$data[, 1:2]
    colnames(df) = c("Axis_1", "Axis_2")
    return(cbind(df, x$data))
})
names(pdataframe)[1] = "method"
```
```{r}
p = ggplot(pdataframe, aes(Axis_1, Axis_2, color=DIAG_CRS, fill=DIAG_CRS))
p = p + geom_point(size=4) #+ geom_polygon()
p = p + facet_wrap(~method, scales="free")
p = p + geom_text(aes(label=subject_id), color = "black")
p
```


## Double Principal Coordinates
```{r, width = 8, height = 7}
# DPCoA Ordination
dpcoa_ordination <- ordinate(filtPhyseqLog, "DPCoA" ) 
dpcoa_evals <- dpcoa_ordination$eig
plot_scree(dpcoa_ordination)
```
```{r}
# Plot DPCoA for samples
dpcoa_ordination_plot_samples <- plot_ordination(filtPhyseqLog, dpcoa_ordination, 
                                                 type = "samples", 
                                                 color = "DIAG_CRS",
                                                 #color = "DIAG_CRS"#, 
                                                 shape = "DIAG_CRS",
                                                 ) +
  geom_text(mapping = aes(label = subject_id), size = 5, color = "red", vjust = 2) +
  geom_point(alpha = 1, size = 3.5) +
  scale_shape_manual(values = c(1,16)) +
  scale_color_manual(values = c("black","black")) +
  theme_classic() +
  theme(
    legend.position = "",
    legend.text = element_text(color = "black"),
    legend.title = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_line(color = "black"),
    axis.text = element_text(color = "black", size = 14),
    axis.title = element_text(color = "black", size = 16),
    panel.border = element_rect(fill = "transparent", color = "black"),
    panel.background = element_rect(fill = "transparent"), # bg of the panel
    plot.background = element_rect(fill = "transparent", color = NA), # bg of the plot
    panel.grid.major = element_line(color = "transparent"), # get rid of major grid
    panel.grid.minor = element_line(color = "transparent"), # get rid of minor grid
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent", color = "transparent") # get rid of legend panel bg
    ) +
  guides(color = guide_legend(override.aes = list(fill="transparent"))) +
  geom_hline(yintercept = 0.0, alpha = 1, linetype = 2, color = "black") +
  geom_vline(xintercept = 0.0, alpha = 1, linetype = 2, color = "black") +
  coord_fixed(sqrt(dpcoa_evals[2] / dpcoa_evals[1])) #+#
  #scale_x_continuous(limits = c(-0.75, .25),
  #                   breaks = c(-0.75,-.5,-.25,0,.25)) +
  #scale_y_continuous(limits = c(-.2,.4),
  #                   breaks = c(-.2,0,.2,.4)) 

# Get rid of point within point
dpcoa_ordination_plot_samples$layers <- dpcoa_ordination_plot_samples$layers[-1]
dpcoa_ordination_plot_samples
```
```{r}
ggsave(dpcoa_ordination_plot_samples, 
       filename = "../figures/ordination/dpcoa-samples.png",
       device = "png",
       height = 3.5,
       width = 5.5)
```

```{r}
OrdLegendSamples <- get_legend(dpcoa_ordination_plot_samples + theme(legend.position = "bottom", legend.title = element_blank()))
ggsave(OrdLegendSamples, filename = "../figures/DPCoASampleLegend.pdf",
       height = 2, width = 5,
       useDingbats = FALSE)
```


```{r}
phylum.colors <- c( "#de2d26", #Actino
                   "#009E73", #Bacteroidetes
                   "#66209d", #Epsiloniarchaeota
                   "#E69F00", #Firmicutes
                   "#F0E442", #Fusobacteria
                   "#D55E00", #Pastesci
                   "#0072B2", #Proteo
                   "#CC79A7",  #Spirochaetes 
                   "purple", #Synergestites
                   "black", #Tenericutes
                   "grey") #Verruco
```

```{r}
# Plot DPCoA scores for taxa
dpcoa_ordination_plot_taxa <- plot_ordination(filtPhyseqLog, dpcoa_ordination, 
                                              type = "taxa", 
                                              color = "Phylum") +
  geom_point(alpha = 1, size = 3.5) +
  scale_color_manual(values = phylum.colors) +
  theme_classic() +
  theme(
    legend.position = "right",
    legend.text = element_text(color = "black"),
    legend.title = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_line(color = "black"),
    axis.text = element_text(color = "black", size = 14),
    axis.title = element_text(color = "black", size = 16),
    panel.border = element_rect(fill = "transparent", color = "black"),
    panel.background = element_rect(fill = "transparent"),
    plot.background = element_rect(fill = "transparent", color = NA), 
    panel.grid.major = element_line(color = "transparent"),
    panel.grid.minor = element_line(color = "transparent"),
    legend.background = element_rect(fill = "transparent"),
    legend.box.background = element_rect(fill = "transparent", color = "transparent"),
    ) +
  guides(#col = guide_legend(ncol = 1),
         color = guide_legend(override.aes = list(fill="transparent"))) +
  geom_hline(yintercept = 0.0, alpha = 1, linetype = 2, color = "black") +
  geom_vline(xintercept = 0.0, alpha = 1, linetype = 2, color = "black") +
  coord_fixed(sqrt(dpcoa_evals[2] / dpcoa_evals[1])) #+
  #scale_x_continuous(limits = c(-0.75, .25),
  #                   breaks = c(-0.75,-.5,-.25,0,.25)) +
  #scale_y_continuous(limits = c(-.2,.4),
  #                   breaks = c(-.2,0,.2,.4))
# Get rid of point within point
dpcoa_ordination_plot_taxa$layers <- dpcoa_ordination_plot_taxa$layers[-1]
dpcoa_ordination_plot_taxa

```

```{r}
OrdLegendTaxa <- get_legend(dpcoa_ordination_plot_taxa + theme(legend.position = "bottom", legend.title = element_blank()))
#ggsave(OrdLegendTaxa, filename = "../figures/DPCoAPhylumLegend.pdf",
#       height = 2, width = 5,
#       useDingbats = FALSE)
```
```{r}
biplot.colors <- c( "black", # Samples
                    "#de2d26", #Actino
                   "#009E73", #Bacteroidetes
                   "#66209d", #Epsiloniarchaeota
                   "#E69F00", #Firmicutes
                   "#F0E442", #Fusobacteria
                   "#D55E00", #Pastesci
                   "#0072B2", #Proteo
                   "#CC79A7",  #Spirochaetes 
                   "purple", #Synergestites
                   "pink", #Tenericutes
                   "grey") #Verruco
```

```{r}
# Plot DPCoA scores for taxa
dpcoa_ordination_biplot <- plot_ordination(filtPhyseqLog, dpcoa_ordination, 
                                              type = "biplot", 
                                              color = "Phylum",
                                           shape = "SAMPLE_TYPE") +
  geom_point(alpha = 1, size = 3) +
  scale_shape_manual(values = c(4,21,19,24)) +
  scale_color_manual(values = biplot.colors) +
  theme_classic() +
  theme(
    legend.position = "right",
    legend.text = element_text(color = "black"),
    legend.title = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_line(color = "black"),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    panel.border = element_rect(fill = "transparent", color = "black"),
plot.margin = unit(c(0,0,0,0), "cm")) +
  guides(#col = guide_legend(ncol = 1),
         color = guide_legend(override.aes = list(fill="transparent"))) +
  geom_hline(yintercept = 0.0, alpha = 1, linetype = 2, color = "black") +
  geom_vline(xintercept = 0.0, alpha = 1, linetype = 2, color = "black") +
  coord_fixed(sqrt(dpcoa_evals[2] / dpcoa_evals[1]))
# Get rid of point within point
dpcoa_ordination_biplot$layers <- dpcoa_ordination_biplot$layers[-1]
dpcoa_ordination_biplot

```
```{r}
ggsave(dpcoa_ordination_biplot, 
       filename = "../figures/ordination/dpcoa-phylum-biplot.pdf",
       device = "pdf",
       height = 4.5,
       width = 8,
       useDingbats = FALSE)
```



```{r}
# Put it all together
dpcoa_split_plot <- cowplot::plot_grid(dpcoa_ordination_plot_samples, dpcoa_ordination_plot_taxa, ncol = 2, align = "h", rel_heights = c(1,1))
dpcoa_split_plot
```


```{r}
# Plot DPCoA for samples
dpcoa_ordination_plot_SNOT20 <- plot_ordination(filtPhyseqLog, dpcoa_ordination, 
                                                 type = "samples", 
                                                 color = "SNOT20_TOTAL",
                                                 #color = "DIAG_CRS"#, 
                                                 shape = "SAMPLE_TYPE"
                                                 ) +
  geom_point(alpha = 0.8, size = 3) +
  theme_classic() +
  coord_fixed(sqrt(dpcoa_evals[2] / dpcoa_evals[1])) +
  scale_color_viridis() +
  scale_shape_manual(values = c(15,16,17)) #For catagorical variables

# Get rid of point within point
dpcoa_ordination_plot_SNOT20$layers <- dpcoa_ordination_plot_SNOT20$layers[-1]
dpcoa_ordination_plot_SNOT20
```
# Ordination - Sinus Only
Just separate out sinus samples for now:

```{r}
FESS <- subset_samples(filtPhyseqRare, SAMPLE_TYPE == "FESS" & SAMPLE_NUMBER == "1") %>% filter_taxa(., function(x) sum(x) > 0, TRUE)
FESS
```
# Transform dataset
```{r}
FESSprop <- transform_sample_counts(FESS, function(x){100 * x / sum(x)})
FESSlog <- transform_sample_counts(FESS, function(x) log(1 + x))
```

Run through some quick ordination plots to look at different ordinations methods
```{r}
dist = "wunifrac"
ord_meths = c("DCA", "CCA", "RDA", "DPCoA", "NMDS", "MDS")
plist = plyr::llply(as.list(ord_meths), function(i, physeq, dist){
        ordi = ordinate(physeq, method=i, distance=dist)
        plot_ordination(physeq, ordi, "samples", color="DIAG_CRS")
}, FESSprop, dist)
```
```{r}
names(plist) <- ord_meths
```
```{r}
pdataframe = plyr::ldply(plist, function(x){
    df = x$data[, 1:2]
    colnames(df) = c("Axis_1", "Axis_2")
    return(cbind(df, x$data))
})
names(pdataframe)[1] = "method"
```
```{r}
p = ggplot(pdataframe, aes(Axis_1, Axis_2, color=DIAG_CRS, fill=DIAG_CRS))
p = p + geom_point(size=4) #+ geom_polygon()
p = p + facet_wrap(~method, scales="free")
p = p + geom_text(aes(label=subject_id), color = "black")
p
```

Look at correlation of the first axis in DPCoA with clinical data:
```{r}
pdataframeDPCoA <- filter(pdataframe, method == "DPCoA")
```

```{r, fig.height=4, fig.width=4}
ggscatter(pdataframeDPCoA, x = "DEM_AGE", y = "Axis_1",
          add = "reg.line",                         # Add regression line
          conf.int = TRUE,                          # Add confidence interval
         # color = "DIAG_CRS",
          palette = "jco",           # Color by groups "cyl"
          shape = "DIAG_CRS"                             # Change point shape by groups "cyl"
          ) +
  stat_cor(#aes(color = DIAG_CRS)
    )           # Add correlation coefficient
```

## wunifrac
```{r}
# DPCoA Ordination
FESS_wuni <- ordinate(FESSprop, 
                            method = "MDS", 
                            distance = "wunifrac") 
FESS_wuni_evals <- FESS_wuni$eig
plot_scree(FESS_wuni)
```

```{r}
# Plot DPCoA scores for taxa
FESS_wuni_mda_splitplot <- plot_ordination(FESSprop, FESS_wuni, type = "split", color = "Phylum", shape = "SAMPLE_TYPE") +
  geom_point(alpha = 1, size = 3) +
  scale_shape_manual(values = c(4,19)) +
  scale_color_manual(values = biplot.colors) +
  theme_pubr(base_size = 10) +
  theme(
    legend.position = "right",
    legend.text = element_text(color = "black"),
    legend.title = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_line(color = "black"),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    panel.border = element_rect(fill = "transparent", color = "black"),
plot.margin = unit(c(0,0,0,0), "cm")) +
  guides(#col = guide_legend(ncol = 1),
         color = guide_legend(override.aes = list(fill="transparent"))) +
  geom_hline(yintercept = 0.0, alpha = 1, linetype = 2, color = "black") +
  geom_vline(xintercept = 0.0, alpha = 1, linetype = 2, color = "black")# +
  #coord_fixed(sqrt(FESS_wuni_evals[2] / FESS_wuni_evals[1]))
# Get rid of point within point
FESS_wuni_mda_splitplot$layers <- FESS_wuni_mda_splitplot$layers[-1]
FESS_wuni_mda_splitplot

```
## Double Principal Coordinates
```{r, width = 8, height = 7}
# DPCoA Ordination
FESS_dpcoa <- ordinate(FESSprop, "DPCoA") 
FESS_dpcoa_evals <- FESS_dpcoa$eig
plot_scree(FESS_dpcoa)
```
```{r}
# Plot DPCoA scores for taxa
FESS_dpcoa_splitplot <- plot_ordination(FESSprop, FESS_dpcoa, type = "split", color = "Order", shape = "SAMPLE_TYPE") +
  geom_point(alpha = .7, size = 3) +
  scale_shape_manual(values = c(20,19)) +
#  scale_color_manual(values = biplot.colors) +
  theme_pubr(base_size = 10) +
  theme(
    legend.position = "right",
    legend.text = element_text(color = "black"),
    legend.title = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_line(color = "black"),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    panel.border = element_rect(fill = "transparent", color = "black"),
plot.margin = unit(c(0,0,0,0), "cm")) +
  guides(#col = guide_legend(ncol = 1),
         color = guide_legend(override.aes = list(fill="transparent"))) +
  geom_hline(yintercept = 0.0, alpha = 1, linetype = 2, color = "black") +
  geom_vline(xintercept = 0.0, alpha = 1, linetype = 2, color = "black") +
  coord_fixed(sqrt(FESS_dpcoa_evals[2] / FESS_dpcoa_evals[1]))
# Get rid of point within point
FESS_dpcoa_splitplot$layers <- FESS_dpcoa_splitplot$layers[-1]
FESS_dpcoa_splitplot
```

```{r}
# Plot DPCoA scores for taxa
FESS_dpcoa_samples <- plot_ordination(FESSprop, FESS_dpcoa, type = "samples") + 
  ggtitle("CF-CRS Samples") +
  geom_point(alpha = .5, shape = 21, size = 3, fill = "black", color = "black") +
  scale_y_continuous(limits = c(-.2,.6),
                     breaks = c(-.2,0,.2,.4,.6)) +
  theme_pubr(base_size = 10) +
  theme(
    plot.title = element_text(size = 12, face = "bold"),
    legend.position = "right",
    legend.text = element_text(color = "black"),
    legend.title = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_line(color = "black"),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    panel.border = element_rect(fill = "transparent", color = "black")
    ) +
  guides(col = guide_legend(ncol = 1),
         color = guide_legend(override.aes = list(fill="transparent"))) +
  geom_hline(yintercept = 0.0, alpha = 1, linetype = 2, color = "black") +
  geom_vline(xintercept = 0.0, alpha = 1, linetype = 2, color = "black") +
  coord_fixed(sqrt(FESS_dpcoa_evals[2] / FESS_dpcoa_evals[1])) 
# Get rid of point within point
FESS_dpcoa_samples$layers <- FESS_dpcoa_samples$layers[-1]
```
```{r}
# Plot DPCoA scores for taxa
FESS_dpcoa_taxa <- plot_ordination(FESSprop, FESS_dpcoa, type = "taxa", color = "Phylum") + 
  ggtitle("Phylum Level Taxa") +
  geom_point(alpha = .5, shape = 20, size = 5) +
  scale_color_manual(values = phylum.colors) +
  scale_y_continuous(limits = c(-.2,.6),
                     breaks = c(-.2,0,.2,.4,.6)) +
  theme_pubr(base_size = 10) +
  theme(
    plot.title = element_text(size = 12, face = "bold"),
    legend.position = "",
    legend.text = element_text(color = "black"),
    legend.title = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_line(color = "black"),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    panel.border = element_rect(fill = "transparent", color = "black")
    ) +
  guides(#col = guide_legend(ncol = 1),
         color = guide_legend(override.aes = list(fill="transparent"))) +
  geom_hline(yintercept = 0.0, alpha = 1, linetype = 2, color = "black") +
  geom_vline(xintercept = 0.0, alpha = 1, linetype = 2, color = "black") +
  coord_fixed(sqrt(FESS_dpcoa_evals[2] / FESS_dpcoa_evals[1])) + 
  ggtitle("Phylum Level Taxa")
# Get rid of point within point
FESS_dpcoa_taxa$layers <- FESS_dpcoa_taxa$layers[-1]
```

```{r}
dpcoa_legend <- get_legend(FESS_dpcoa_taxa + theme(legend.position = "right"))
ggsave(dpcoa_legend, filename = "../figures/ordination/dpcoa_phylum_legend.pdf",
       height = 3, width = 2)
```


```{r, fig.height=7, fig.width=5}
dpcoa_samples_taxa <- plot_grid(FESS_dpcoa_samples, FESS_dpcoa_taxa,
                                ncol = 1, align = "v")
dpcoa_samples_taxa
```
```{r}
ggsave(dpcoa_samples_taxa, filename = "../figures/ordination/dpcoa-samples-phylum.pdf",
       device = "pdf",
       height = 5,
       width = 4,
       useDingbats = FALSE)
```

## Look at Sinus Factors
For DF508 we have to recalc distances
```{r}
# Remove sample that is NA for genotype data
FESSpropNo87 <- subset_samples(FESSprop, !subject_id == "87")  %>% filter_taxa(., function(x) sum(x) > 0, TRUE)

# Get dataframe of sample data for this dataset
sampledfNo87 <- data.frame(sample_data(FESSpropNo87))
# Calculate distance matrix
dpcoaNo87 <- phyloseq::distance(FESSpropNo87, method = "dpcoa")
# Adonis test - dF508
adonis(dpcoaNo87 ~ HOMO_dF508, data = sampledfNo87)
# Calculate ordination to plot
dpcoaNo87ord <- ordinate(FESSpropNo87, "DPCoA") 
# Plot Ordination
ord <- plot_ordination(FESSpropNo87, dpcoaNo87ord, type = "samples", color = "HOMO_dF508")
ord
```


```{r fig.width=10, fig.height=10}
cat_colors <- c("dodgerblue","red", "grey")
theme_set(theme_bw())
# Plot DPCoA scores for categorical variables
# SINUS FACTORS
FESS_dpcoa_samples_polyps <- plot_ordination(FESSprop, FESS_dpcoa, type = "samples", color = as.character("SINUS_FACTORS_POLYPS")) + geom_point(size = 3, alpha = 0.75) + theme(legend.title = element_blank()) + scale_color_manual(values = cat_colors) + theme(legend.position = "") + ggtitle("Polyps")
FESS_dpcoa_samples_asthma <- plot_ordination(FESSprop, FESS_dpcoa, type = "samples", color = as.character("SINUS_FACTORS_ASTHMA")) + geom_point(size = 3, alpha = 0.75) + theme(legend.position = "") + scale_color_manual(values = cat_colors) + ggtitle("Asthma")
FESS_dpcoa_samples_gerd <- plot_ordination(FESSprop, FESS_dpcoa, type = "samples", color = as.character("SINUS_FACTORS_GERD")) + geom_point(size = 3, alpha = 0.75) + theme(legend.position = "") + scale_color_manual(values = cat_colors) + ggtitle("GERD")
FESS_dpcoa_samples_allergies <- plot_ordination(FESSprop, FESS_dpcoa, type = "samples", color = as.character("SINUS_FACTORS_ALLERGIES")) + geom_point(size = 3, alpha = 0.75) + theme(legend.position = "") + scale_color_manual(values = cat_colors) + ggtitle("Allergies")
FESS_dpcoa_samples_priorfess <- plot_ordination(FESSprop, FESS_dpcoa, type = "samples", color = as.character("SINUS_FACTORS_PRIORFESS")) + geom_point(size = 3, alpha = 0.75) + theme(legend.position = "") + scale_color_manual(values = cat_colors) + ggtitle("Prior FESS")
FESS_dpcoa_samples_dF508 <- plot_ordination(FESSpropNo87, dpcoaNo87ord, type = "samples", color = as.character("HOMO_dF508")) + geom_point(size = 3, alpha = 0.75) + scale_color_manual(values = cat_colors) + ggtitle("DF508")

# Remove point in first layer for aesthetic purposes
FESS_dpcoa_samples_polyps$layers <- FESS_dpcoa_samples_polyps$layers[-1]
FESS_dpcoa_samples_asthma$layers <- FESS_dpcoa_samples_asthma$layers[-1]
FESS_dpcoa_samples_gerd$layers <- FESS_dpcoa_samples_gerd$layers[-1]
FESS_dpcoa_samples_allergies$layers <- FESS_dpcoa_samples_allergies$layers[-1]
FESS_dpcoa_samples_priorfess$layers <- FESS_dpcoa_samples_priorfess$layers[-1]
FESS_dpcoa_samples_dF508$layers <- FESS_dpcoa_samples_dF508$layers[-1]
# Plot grid
FESS_dpcoa_samples_legend <- get_legend(FESS_dpcoa_samples_DF508)
FESS_dpcoa_sinusfactors <- plot_grid(FESS_dpcoa_samples_polyps,
                                     FESS_dpcoa_samples_asthma,FESS_dpcoa_samples_gerd, FESS_dpcoa_samples_allergies, 
                                     FESS_dpcoa_samples_priorfess, 
                                     FESS_dpcoa_samples_dF508 + theme(legend.position = ""),
                                     nrow = 3, ncol = 3)
FESS_dpcoa_sinusfactors
FESS_dpcoa_samples_legend <- plot_grid(FESS_dpcoa_samples_legend)
FESS_dpcoa_samples_legend
```
```{r}
ggsave(FESS_dpcoa_sinusfactors, 
       filename = "../figures/ordination/dpcoa-sinus-factors.pdf",
       device = "pdf",
       height = 7,
       width = 7,
       useDingbats = FALSE)
```


```{r}
set.seed(1)

# Calculate bray curtis distance matrix
dpcoa <- phyloseq::distance(FESSprop, method = "dpcoa")

# make a data frame from the sample_data
sampledf <- data.frame(sample_data(FESSprop))

# Adonis test
adonis(dpcoa ~ SINUS_FACTORS_POLYPS, data = sampledf)

# Adonis test - asthma
adonis(dpcoa ~ SINUS_FACTORS_ASTHMA, data = sampledf)

# Adonis test - allergies
adonis(dpcoa ~ SINUS_FACTORS_ALLERGIES, data = sampledf)

# Adonis test - GERD
adonis(dpcoa ~ SINUS_FACTORS_GERD, data = sampledf)
# Adonis test - Number of FESS
adonis(dpcoa ~ SINUS_FACTORS_PRIORFESS, data = sampledf)





```



```{r}
# Plot DPCoA scores for taxa
FESS_dpcoa_samples_snot20 <- plot_ordination(FESSprop, FESS_dpcoa, type = "samples", color = "DEM_AGE") +
  geom_point(alpha = 1, size = 3) +
  theme_classic() +
  theme(
    legend.position = "right",
    legend.text = element_text(color = "black"),
    legend.title = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_line(color = "black"),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    panel.border = element_rect(fill = "transparent", color = "black"),
plot.margin = unit(c(0,0,0,0), "cm")) +
  guides(#col = guide_legend(ncol = 1),
         color = guide_legend(override.aes = list(fill="transparent"))) +
  geom_hline(yintercept = 0.0, alpha = 1, linetype = 2, color = "black")
  geom_vline(xintercept = 0.0, alpha = 1, linetype = 2, color = "black") #+
  #coord_fixed(sqrt(FESS_dpcoa_evals[2] / FESS_dpcoa_evals[1]))
# Get rid of point within point
#FESS_dpcoa_samples_snot20$layers <- FESS_dpcoa_samples_snot20$layers[-1]
FESS_dpcoa_samples_snot20
```
correlation with the first axis and age?





Notes:

Comparisons to make:
- Look at genotype? OR class. Differences in dF508 vs heterogeneous.
- Look at SNOT20 scores when get it.
- Do ordinations without sputum samples.
